Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-03-31 main.tex
glued up flattening
minor changes

% Pretty references
\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{th}{Theorem~\ref{#1}}
\newrefformat{co}{Corollary~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{cr}{Condition~\ref{#1}}
\newrefformat{crs}{Conditions~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

@@

% Macros
\newcommand{\sN}{\mathbb{N}}
\newcommand{\sNN}{\mathbb{N}^{**}}
\def\DEF{\stackrel{\Delta}=}
\def\EQDEF{\stackrel{\Delta}\Leftrightarrow}
\newcommand{\segm}[2]{\llbracket #1; #2 \rrbracket}
\def\indexes#1{\mathbb{I}^{#1}}
\def\f#1{\mathsf{#1}}
\def\prio{\mathsf{prio}}
\newcommand{\bottom}{\perp}
\newcommand{\stable}{\mathsf{stable}}
\newcommand{\update}{\mathsf{update}}
\newcommand{\components}{\Gamma}
%\newcommand{\cs}{CS}
\newcommand{\cs}{\Delta}

@@

To address this issue, we propose an extended definition of the framework, including priority classes for actions.
In this paper, we focus on a restriction of the Process Hitting with two classes of priorities and a specific behaviour of the components, that is sufficient to tackle the aforementioned problem of cooperations.
We show that this class of Process Hitting models allows to represent any
Asynchronous Discrete Networks, either Boolean or multivalued; and also any
Process Hitting model with arbirtrary number of priorty classes.
Then we develop a new refinement for the under-approximation of the static analysis to give accurate results for this class of Process Hitting models.
Our method thus allows to efficiently under-approximate reachability properties in Asynchronous Discrete Networks;
it is in particular conclusive on reachability properties in a 94 components Boolean network, which is unprecedented.
\end{abstract}
\begin{keyword}
discrete networks \sep
abstract interpretation \sep
reachability \sep
qualitative models \sep
systems biology
\end{keyword}
\end{frontmatter}

>>>>

% Pretty references
\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{th}{Theorem~\ref{#1}}
\newrefformat{co}{Corollary~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{cr}{Condition~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

@@

% Macros
\newcommand{\sN}{\mathbb{N}}
\newcommand{\sNN}{\mathbb{N}^\bullet}
\def\DEF{\stackrel{\Delta}=}
\def\EQDEF{\stackrel{\Delta}\Leftrightarrow}
\newcommand{\segm}[2]{\llbracket #1; #2 \rrbracket}
\def\indexes#1{\mathbb{I}^{#1}}
\def\f#1{\mathsf{#1}}
\def\prio{\mathsf{prio}}
\newcommand{\bottom}{\perp}
\newcommand{\stable}{\mathsf{stable}}
\newcommand{\update}{\mathsf{update}}
\newcommand{\components}{\Gamma}
%\newcommand{\cs}{CS}
\newcommand{\cs}{\Delta}

@@

To address this issue, we propose an extended definition of the framework, including priority classes for actions.
In this paper, we focus on a restriction of the Process Hitting with two classes of priorities and a specific behaviour of the components, that is sufficient to tackle the aforementioned problem of cooperations.
We show that this class of Process Hitting models allows to represent any
Asynchronous Discrete Networks, either Boolean or multivalued; and also any
Process Hitting model with arbirtrary number of priorty classes.
Then we develop a new refinement for the under-approximation of the static analysis to give accurate results for this class of Process Hitting models.
Our method thus allows to efficiently under-approximate reachability properties
in Asynchronous Discrete Networks,
as it is illustrated on the model-checking of a signaling network of 94
components, which is unprecedented.
\end{abstract}
\begin{keyword}
discrete networks \sep
abstract interpretation \sep
reachability \sep
qualitative models \sep
systems biology
\end{keyword}
\end{frontmatter}

### 2014-03-31 parts/intro.tex
minor changes

This paper is an extension of \cite{FPMR13-CS2Bio} were we
we introduced a new extension to the semantics of Process Hitting by partitioning actions into classes of priorities
(\pref{sec:ph}).
One of the objectives is to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability of the state of components on a subclass of this new framework, thus allowing to compute efficient static analysis.
This local reachability under-approximation can be also easily extended to study the reachability of a global state.
Furthermore, we provide a new extended under-approximation method to decide a successive reachability property.

Whereas the analysis presented in \pref{sec:sa} considers a particular class of
Process Hitting with two classes of priority,
we provide in \pref{sec:encodings} formal encodings of Asynchronous Discrete Networks and general
Process Hitting models with priorities.
Those encodings result in a Process Hitting fulfilling the conditions for our
static analysis and show the broad applicability of our method.

>>>>

This paper is an extension of \cite{FPMR13-CS2Bio} were we
introduced a new extension to the semantics of Process Hitting by partitioning actions into classes of priorities
(\pref{sec:ph}).
One of the objectives is to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability of the state of components on a subclass of this new framework, thus allowing to compute efficient static analysis.
This local reachability under-approximation can be also easily extended to study the reachability of a global state.
Furthermore, we provide a new extended under-approximation method to decide a successive reachability property.

Whereas the analysis presented in \pref{sec:sa} considers a particular class of
Process Hitting with two classes of priority,
we provide in \pref{sec:encodings} formal encodings of Asynchronous Discrete Networks and general
Process Hitting models with priorities.
Those encodings result in a Process Hitting fulfilling the conditions for our
static analysis and support the broad applicability of our method.

### 2014-03-31 parts/sa.tex
minor changes

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in \PHl, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in \PHl, s\subset\ctx$, there exists a scenario
concretizing $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}

>>>>

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a Process Hitting $(\PHs; \PHl; \PHa^{\langle 2 \rangle})$
that satisfies \pref{cr:bounded} and \pref{cr:compcs},
a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in \PHl, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in \PHl, s\subset\ctx$, there exists a scenario
concretizing $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}
